Incentives are the biggest barriers to widespread data sharing

### 2021-08-13 new-paper.Rmd
minor changes to first two pages

---
title: Incentives are the biggest barriers to widespread data sharing
author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
citation_package: 'natbib'
csl: nature-communications.csl
header-includes:
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

```{=tex}
\begin{singlespace}
1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia\\
2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
3 = Telethon Kids Institute, Perth Children's Hospital, Perth, Australia\\
4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
* = Both authors contributed equally to the work\\
Correspondence to:\\
Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (karthik.ram@berkeley.edu)

@@

```{=tex}
\begin{singlespace}
There are numerous arguments that strongly support the practice of open science, with societal and individual benefits. For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations. Despite the general appreciation of the benefits of data sharing, research data are often only available to the original investigators. For data that are shared, lack of useful metadata and documentation make them challenging to reuse. In this paper we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing. We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap. While previous papers have focused on describing ideal best practices for data and code, we focus on  common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.

@@

> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)

The idea of open and transparent science dates back to the founding principles of The Royal Society of London in 1660. The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] primarily among early career researchers [@McKiernan2016]. A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings. Jon Clarebout famously wrote in 1992 that reproducibility in computationally oriented fields had become so trivial that non experts could do the task. Clarebout's optimistic vision was a little ahead of its time as the tools and technologies available in the early nineties for literate programming presented too high of a barrier for most researchers to adopt without incurring a significant loss in productivity [@Barnes2010].

Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence. Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015]. The bigger driver for this is more likely due to the growing popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools. Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data. Each of these languages enjoys the support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. Modern scientists now have access to many interactive computational frameworks (e.g. Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages. The barriers to verification are now low enough that reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].

Notebooks however, are an incomplete solution for computational reproducibility. For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code; 2) Computing environment, and 3) Data. There has been tremendous progress on the first two, despite some limitations of existing tools and practices. While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations played a strong role in the culture change. Sharing code wasn't just easy at first [@Barnes2010], but became a social and normative part of academic culture. The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014], tooling, and recognition as a scholarly contribution [@Smith2018]. Getting researchers to share usable data has been more of a struggle.

>>>>

---
title: Incentives are the biggest barriers to widespread data sharing
author: Nicholas J. Tierney\* (1, 2, 3) & Karthik Ram\* (4)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
citation_package: 'natbib'
csl: nature-communications.csl
header-includes:
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

```{=tex}
\begin{singlespace}
1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia\\
2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
3 = Telethon Kids Institute, Perth Children's Hospital, Perth, Australia\\
4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
* = Both authors contributed equally to the work.\\
Correspondence to:\\
Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (karthik.ram@berkeley.edu)

@@

```{=tex}
\begin{singlespace}
Numerous arguments strongly support the practice of open science, which offers several societal and individual benefits. For individual researchers, sharing research artifacts such as data can increase trust and transparency, improve the reproducibility of one's own work, and catalyze new collaborations. Despite a general appreciation of the benefits of data sharing, research data are often only available to the original investigators. For data that are shared, lack of useful metadata and documentation make them challenging to reuse. In this paper, we argue that a lack of incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing. We compare data to code, examine computational environments in the context of  their ability to facilitate the reproducibility of research, provide some practical guidance on how one can improve the chances of their data being reusable, and partially bridge the incentive gap. While previous papers have focused on describing ideal best practices for data and code, we focus on  common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.

@@

> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." -- Sherlock Holmes ("The Adventure of the Copper Beeches" by Sir Arthur Conan Doyle)

The idea of open and transparent science dates back to the founding principles of The Royal Society of London in 1660. The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] primarily among early career researchers [@McKiernan2016]. A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust in and verification of those findings. Jon Clarebout famously wrote in 1992 that reproducibility in computationally-oriented fields had become so trivial that nonexperts could do the task. Clarebout's optimistic vision was a little ahead of its time, as the tools and technologies available in the early 1990s for literate programming presented too high a barrier for most researchers to adopt without incurring a significant loss in productivity [@Barnes2010].

Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables, and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence. Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015]. The bigger driver for this is more likely the growing popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open-source software tools. Programming languages such as Python, R, and Julia have helped scientists implement and share new methods for working with data. Each of these languages enjoys the support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. Modern scientists now have access to many interactive computational frameworks (e.g. Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages. The barriers to verification are now low enough that reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].

Notebooks, however, are an incomplete solution for computational reproducibility. For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) code; 2) computing environment, and 3) data. There has been tremendous progress on the first two, despite some limitations of existing tools and practices. While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations have played a strong role in the culture change. Sharing code wasn't easy at first [@Barnes2010], but became a social and normative part of academic culture. The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014], tooling, and recognition of code as a scholarly contribution [@Smith2018]. Getting researchers to share usable data has been more of a struggle.
